---
title: "Data Classes - Exercises"
author: "Gbemisola Talabi"
date: "21 September 2022"
output: pdf_document
---


# Part 1

Load all the libraries we will use in this lab. 

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(lubridate)
```

0. Write you name in the YAML header above

1. Create some data to work with. 

First, create a vector that has class integer called `int_vect` that starts at 1 and goes up to 10 and repeats this sequence 3 times using `rep`  (hint use `seq()`).

```{r}
int_vect <- as.integer(rep(seq(from =1, to = 10, by = 1), times =3))
is.integer(int_vect)
class(int_vect)
```


2. Next, create a random vector of 30 values from a set of integers from 1 to 30 called `rand_vect` (hint use `sample()` and set the replace argument to `TRUE`).

Because we are using a random vector let's use the `set.seed()` function to make sure we all have the same result - this code is already in the code chunk for you. Simply create the vector below the `set.seed` line.


```{r}
set.seed(1234)
rand_vect <- sample(
  x = seq(from = 1, to = 30, by = 1),
  size = 30, replace = TRUE
)

# this also works... need to set the seed right before
set.seed(1234)
rand_vect <- sample(x = 1:30, size = 30, replace = TRUE)
```

3a. Create a vector that repeats `c(TRUE, TRUE, FALSE)` 10 times called `TF_vect`. Also create a vector that repeats `c("TRUE", "TRUE", "FALSE")` 10 times called `TF_vect2`.

```{r}
TF_vect <- rep(c(TRUE, TRUE, FALSE), times =10)
TF_vect2 <- rep(c("TRUE", "TRUE", "FALSE"), times = 10)

```

3. Create a tibble combining these vectors together called `vect_data` using the following code.

```{r}
vect_data <- tibble(int_vect, rand_vect, TF_vect, TF_vect2)
```


4. Take a look at 5 random rows using the `slice_sample()` function. Try this a few times to see how the results change.

```{r}
slice_sample(vect_data, n = 5)

```

5. Take a look at 5 random rows using the `slice_sample()` function again but this time with `set.seed(1234)` as the first line of the chunk. Try this a few times to see the results. (Don't forget to not copy the backticks.)

```{r}
set.seed(1234)
slice_sample(vect_data, n = 5)
```

6. Check to see if the `TF_vect` is logical. Check to see if `TF_vect2` is logical. Why are the results what they are?

```{r}
class(TF_vect)
class(TF_vect2)
# because TF_vect is not in quotes, it is a logical element.
#because TF_vect2 is in quotes, it is a character.

```


7. Use `mutate()` function to create a new variable in `vect_data` named `type_fact` that is of class `factor` made from the `int_vect` variable. Take a look at the data and observe how the class is different for the new variable compared to `int_vect`. 

```{r}
vect_data <- vect_data %>% mutate(type_fact = as.factor(int_vect))

```


# Part 2

1. Read in the Charm City Circulator data file "circulator_ridership.csv". We have used this data before and should be in your data folder, but you can also download in again from Day1. Remember that a .RMD file automatically makes its working directory where the .RMD file is located.  So you will likely need to use `../data/circulator_ridership.csv` as the relative path to move out the subdirector with `../` and then back into the data folder with `data/` 
  
```{r}
charmcity<- read_csv("../data/circulator_ridership.csv")
```


Use the `str()` function to take a look at the data and learn about the column types.

```{r}
str(charmcity)
```

2. Use the `mutate()` function to create a new column named `date_formatted` that is of `Date` class. The new variable is created from `date` column. Hint: use `mdy()` function.

```{r}

charmcity <- charmcity %>% mutate( 
              date_formatted = ymd(date),
              date = as.numeric(date))
glimpse(charmcity)

```

3. Move the `date_formatted` variable to be before `date` using the `relocate` function. Take a look at the data using `glimpse()`. Note the difference between `date` and `date_formatted` columns.  

```{r}
charmcity <- charmcity %>% relocate(date_formatted, .before = date)
glimpse(charmcity)

```

4. Create a list data object called `classes_data` that combines the `vect_data` from the first part of the lab and `circ`. Use `glimpse()` to look at the data.

```{r}
classes_data <- list(vect_data, charmcity)
glimpse(classes_data)

```

5. Use `range()` function on `date_formatted` variable to display the range of dates in the data set. How does this compare to that of `date`? Why? (Hint: use the pull function first to pull the values.)

```{r}
range(charmcity$date_formatted)
range(charmcity$date)
#the date_formatted is in date class(year/month/day) and the date is in numerical class. 
#because we changed the date_formatted variable to Date class
```


6. Use the `group_by` function on `day` and `line` variables with the `summarize()` function, to display the `number` of orange boardings observations with each day (hint: `use sum()`). Which day had the most boardings? Is this true for the other routes (purple boardings,  green boardings,  banner boardings)?

```{r}
out <- charmcity %>% 
  filter(type=="Boardings") %>%
  group_by(day, line) %>%
  summarise(sum_var = sum(number, na.rm=T)) %>%

sum((charmcity$line == "orange") & (charmcity$type == "Boardings"))
sum((charmcity$line == "purple") & (charmcity$type == "Boardings"))
sum((charmcity$line == "green") & (charmcity$type == "Boardings"))
sum((charmcity$line == "banner") & (charmcity$type == "Boardings"))

# yes, this is true for every other route
```
 

